---
title: "Getting Started with rdtp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rdtp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

## Overview

`rdtp` searches over candidate cutoff values on a discrete forcing (running)
variable to find, for each unit, the cutoff that maximizes R-squared from a
linear-spline regression discontinuity (RD) regression. This is the first step
of a "tipping-point" analysis: identifying where each unit (e.g., school,
county, hospital) appears to impose a threshold on a test score or other
discrete assignment variable.

The method is based on McEachin, Domina, and Penner (2020, *JPAM*). See also
Hansen (2000, *Econometrica*) for the econometric foundations of threshold
detection.

## Installation

```{r install, eval = FALSE}
# Install from GitHub (when available):
# remotes::install_github("amceachin-code/rdtp")

# Or install from local source:
# install.packages("path/to/rdtp", repos = NULL, type = "source")
```

## Synthetic Data Example

We generate data for 5 schools, each with 100 students and a different true
cutoff on a discrete test-score variable. A positive discontinuity of +20
points is applied at each school's cutoff.

```{r generate-data}
library(rdtp)

set.seed(12345)
n <- 500

dat <- data.frame(
  school     = rep(1:5, each = 100),
  test_score = 200 + sample(0:300, n, replace = TRUE)
)

# True cutoffs: 300, 350, 400, 325, 375
true_cuts <- c(300, 350, 400, 325, 375)
dat$above <- as.integer(dat$test_score >= true_cuts[dat$school])
dat$outcome <- 50 + 0.1 * dat$test_score + 20 * dat$above + rnorm(n, 0, 3)
```

## Running the Search

```{r basic-run}
fit <- rdtp(dat,
            depvar      = "outcome",
            forcing     = "test_score",
            by          = "school",
            searchrange = c(250, 450))
```

## Viewing Results

```{r print}
print(fit)
```

The `summary()` method provides distributional statistics:

```{r summary}
summary(fit)
```

## Plotting

Three plot types are available: `"cutoffs"`, `"r2"`, and `"beta"`.

```{r plot-cutoffs}
plot(fit, type = "cutoffs")
```

```{r plot-r2}
plot(fit, type = "r2")
```

```{r plot-beta}
plot(fit, type = "beta")
```

## Options

### Bandwidth

The `bandwidth` parameter sets the symmetric bandwidth around each candidate
cutoff. Only observations with `|forcing - c| <= bandwidth` enter the
regression. Default is 75.

```{r bandwidth}
fit_narrow <- rdtp(dat,
                   depvar    = "outcome",
                   forcing   = "test_score",
                   by        = "school",
                   bandwidth = 50,
                   searchrange = c(250, 450))
fit_narrow
```

### Robust Standard Errors

Use `vce = "robust"` for heteroskedasticity-robust (HC1) standard errors.
Requires the `sandwich` package.

```{r robust, eval = requireNamespace("sandwich", quietly = TRUE)}
fit_rob <- rdtp(dat,
                depvar  = "outcome",
                forcing = "test_score",
                by      = "school",
                vce     = "robust",
                searchrange = c(250, 450))
fit_rob$results[, c("unit", "beta", "se")]
```

### Cluster-Robust Standard Errors

Use `vce = "cluster"` with the `cluster` argument. Requires the `sandwich`
package.

```{r cluster, eval = requireNamespace("sandwich", quietly = TRUE)}
dat$classroom <- rep(1:25, each = 20)

fit_cl <- rdtp(dat,
               depvar  = "outcome",
               forcing = "test_score",
               by      = "school",
               vce     = "cluster",
               cluster = "classroom",
               searchrange = c(250, 450))
fit_cl$results[, c("unit", "beta", "se")]
```

### Saving Results

Save results to `.rds` or `.csv`:

```{r saving, eval = FALSE}
fit <- rdtp(dat,
            depvar  = "outcome",
            forcing = "test_score",
            by      = "school",
            saving  = "my_results.rds")

# Or CSV:
fit <- rdtp(dat,
            depvar  = "outcome",
            forcing = "test_score",
            by      = "school",
            saving  = "my_results.csv")
```

### Verbose Output

Set `verbose = TRUE` to see per-unit progress:

```{r verbose, eval = FALSE}
fit <- rdtp(dat,
            depvar  = "outcome",
            forcing = "test_score",
            by      = "school",
            verbose = TRUE,
            searchrange = c(250, 450))
```

## References

- McEachin, A., T. Domina, and A. Penner. 2020. "Heterogeneous effects of
  early algebra across California middle schools." *Journal of Policy Analysis
  and Management* 39(3): 772-800.
  [doi:10.1002/pam.22202](https://doi.org/10.1002/pam.22202)

- Hansen, B. E. 2000. "Sample splitting and threshold estimation."
  *Econometrica* 68(3): 575-603.
  [doi:10.1111/1468-0262.00124](https://doi.org/10.1111/1468-0262.00124)

- Lee, D. S., and T. Lemieux. 2010. "Regression discontinuity designs in
  economics." *Journal of Economic Literature* 48(2): 281-355.
  [doi:10.1257/jel.48.2.281](https://doi.org/10.1257/jel.48.2.281)
